---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r opts, echo = FALSE}
library(nlstimedist)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

```

# nlstimedist

[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![Travis-CI Build Status](https://travis-ci.org/NattyE/nlstimedist.svg?branch=master)](https://travis-ci.org/NattyE/nlstimedist)
[![codecov](https://codecov.io/gh/NattyE/nlstimedist/branch/master/graph/badge.svg)](https://codecov.io/gh/NattyE/nlstimedist)

nlstimedist fits a biologically meaningful distribution function to time-sequence data (phenology), estimates parameters to draw the cumulative distribution function and probability density function and calculates standard statistical moments and percentiles.

# Installation

You can install:

* the latest released version from CRAN with

```{r cran, eval = FALSE}
install.packages("nlstimedist")
```

* the latest development version from GitHub with

```{r gh, eval = FALSE}
devtools::install_github("NattyE/nlstimedist")
```

# Usage

## Preparing the data
Data should be in tidy format. `nlstimedist` provides three example tidy datasets: `lobelia`, `pupae` and `tilia`.

```{r data}
head(tilia)
```

We first need to calculate the cumulative number of trees as well as the proportions. We do this using the `tdData` function.

```{r tidy}
tdTilia <- tdData(tilia, x = "Day", y = "Trees")
tdTilia
```

## Fitting the model
We fit the model to the proportion of the cumulative number of trees (`propMax`) in the `tdTilia` data using the `timedist` function.

```{r model}
model <- timedist(data = tdTilia, x = "Day", y = "propMax", r = 0.1, c = 0.5, t = 120)
model
```

## Extracting the moments
We can extract the mean, variance, standard deviation, skew, kurtosis and entropy of the model as follows.

```{r moments}
model$m$getMoments()
```

## Extracting the RSS
Similarly we can extract the RSS of the model

```{r rss}
model$m$rss()
```

## Plotting the PDF and CDF
The pdf and cdf of the model have their own plotting functions.

```{r pdfPlot}
tdPdfPlot(model)
```

```{r cdfPlot}
tdCdfPlot(model)
```

# Citation
Franco, M. (2012). *The time-course of biological phenomenon - illustrated with the London Marathon*. Unpublished manuscript. Plymouth University.
