---
title: "Package Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(nlstimedist)
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```
  
# Fitting the time-course of the leafing phenology of lime trees

## Preparing the data
Data should be in tidy format. `nlstimedist` provides three example tidy datasets: `lobelia`, `pupae` and `tilia`.

```{r}
head(tilia)
```

We first need to calculate the cumulative number of trees as well as the proportions. We do this using the `tdData` function.

```{r}
tdTilia <- tdData(tilia, x = "Day", y = "Trees")
tdTilia
```

## Fitting the model
We fit the model to the proportion of the cumulative number of trees (`propMax`) in the `tdTilia` data using the `timedist` function.

```{r}
model <- timedist(data = tdTilia, x = "Day", y = "propMax", r = 0.1, c = 0.5, t = 120)
model
```

## Extracting the moments
We can extract the mean, variance, standard deviation, skew, kurtosis and entropy of the model as follows.

```{r}
model$m$getMoments()
```

## Extracting the RSS
Similarly we can extract the RSS of the model

```{r}
model$m$rss()
```

## Plotting the PDF and CDF
The pdf and cdf of the model have their own plotting functions.

```{r}
tdPdfPlot(model)
```

```{r}
tdCdfPlot(model)
```
